Hierarchy Builder: algebraic~hierarchies made easy in Coq with Elpi

### 2020-04-24 hb.tex
Address the =KS= items except explanation of the flat variant

The language of packed classes were introduced in
\cite{DBLP:conf/tphol/GarillotGMR09} and used directly to describe the
algebraic hierarchy of the Mathematical Components library.
It is based on a disciplined use of records and projections and
on the possibility of extending the elaborator of Coq
via the declaration of Canonical Structures instances.
In this section we describe the \emph{flat} variant of
Packed Classes, the target language of \HB{}.

@@

Next we define Abelian groups.
Since the monoid structure is the bottom of this hierarchy its class record
\coq{Monoid} consists just one \mixin{}. In contrast the class record
of Abelian groups consists of two \mixins{} where the second one depends
 on the former (since Abelian groups inherit from monoids).
\begin{coqcode}
Module AbelianGroup.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type T;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort }.
End AbelianGroup.
\end{coqcode}
As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
coercion and lift additive inverse \coq{opp} and Abelian group axioms.
Since Abelian groups inherit from monoids, we also declare an implicit coercion
from Abelian groups to monoids. A coercion between structures can be defined
in two steps: first a coercion between the class record of the
superclass to the class record of the subclass (line~\ref{demo4:coe:ag:monoid:class});
and a coercion between
structure records (line~\ref{demo4:coe:ag:monoid:type}) relying on the first one.
\begin{coqcode}
Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup T) : #\label{demo4:coe:ag:monoid:class}#
  Monoid T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
  Monoid.Pack (AbelianGroup.sort T) (AbelianGroup.class T).

@@

\subsection{Multiple inheritance}
In order to introduce multiple inheritance, we extend the hierarchy described in
section~\ref{subsec:v3} with the structure of semirings%, which is another
%intermediate structure between monoids and rings 
as depicted in figure \ref{fig:puzzle_v4}. Semirings introduce the binary
multiplication operator \coq{mul} and its identity element \coq{one}.

@@

We recall here how to use the Canonical
Structures~\cite{DBLP:conf/itp/MahboubiT13,Saibi99phd} mechanism, which
lets the user extend the elaborator of Coq, in order to handle
inheritance and inference of structure~\cite{affeldt:hal-02463336,KSdraft,DBLP:conf/tphol/GarillotGMR09}.
This software component
takes in input a term as written by the user and has to infer all the missing information that is necessary in order to make the term well typed.

>>>>

The language of Packed Classes~\cite{DBLP:conf/tphol/GarillotGMR09} is used directly to describe the algebraic hierarchy of the Mathematical Components library.
It is based on a disciplined use of records and projections and on the possibility of extending the elaborator of Coq via the declaration of Canonical Structures~\cite{DBLP:conf/itp/MahboubiT13} instances.
In this section, we describe the \emph{flat} variant of Packed Classes, the target language of \hb{}, through the hierarchy V3 extended with semirings.

The \hb{} addon can generate all the Coq declarations in this section automatically, but some details are omitted for brevity in this section.

@@

Next we define Abelian groups.
Since the monoid structure is the bottom of this hierarchy its class record
\coq{Monoid} consists of just one \mixin{}. In contrast the class record
of Abelian groups consists of two \mixins{} where the second one depends
 on the former (since Abelian groups inherit from monoids).
\begin{coqcode}
Module AbelianGroup.
Record axioms (T : Type) : Type := Class {
  Monoid_of_Type_mixin : Monoid_of_Type T;
  AbelianGroup_of_Monoid_mixin : AbelianGroup_of_Monoid T Monoid_of_Type_mixin; }.
Structure type : Type := Pack { sort : Type; class : axioms sort }.
End AbelianGroup.
\end{coqcode}
As in the monoid structure, we declare \coq{AbelianGroup.sort} as an implicit
coercion and lift additive inverse \coq{opp} and Abelian group axioms.
Since Abelian groups inherit from monoids, we also declare an implicit coercion
from Abelian groups to monoids. A coercion between structures can be defined
in two steps: first a coercion between the class record of the
superclass to the class record of the subclass (line~\ref{demo4:coe:ag:monoid:class});
and a coercion between
structure records (line~\ref{demo4:coe:ag:monoid:type}) relying on the first one.
\begin{coqcode}
Coercion AbelianGroup.sort : AbelianGroup.type >-> Sortclass.
Coercion AbelianGroup_class_to_Monoid_class (T : Type) (c : AbelianGroup T) : #\label{demo4:coe:ag:monoid:class}#
  Monoid T := Monoid.Class T (AbelianGroup.Monoid_of_Type_mixin T c).
Coercion AbelianGroup_to_Monoid (T : AbelianGroup.type) : Monoid.type := #\label{demo4:coe:ag:monoid:type}#
  Monoid.Pack (AbelianGroup.sort T) (AbelianGroup.class T).

@@

\subsection{Multiple inheritance}
In order to introduce multiple inheritance, we extend the hierarchy described in
section~\ref{subsec:v3} with the structure of semirings %, which is another
%intermediate structure between monoids and rings 
as depicted in figure \ref{fig:puzzle_v4}. Semirings introduce the binary
multiplication operator \coq{mul} and its identity element \coq{one}.

@@

We recall here how to use the Canonical
Structures~\cite{DBLP:conf/itp/MahboubiT13,Saibi99phd} mechanism, which
lets the user extend the elaborator of Coq, in order to handle
inheritance and inference of structure~\cite{affeldt:hal-02463336,KSdraft,DBLP:conf/tphol/GarillotGMR09}.
This software component
takes as input a term as written by the user and has to infer all the missing information that is necessary in order to make the term well typed.
